### 2020-12-20 paper/abstract.tex
Move to Holger's templates and some other - cannot remember - fixes.

\begin{abstract}

Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases while also  minimizing social and economic costs. Epidemiological models have a hard time predicting the effects of such fine-grained policies. We propose a novel simulation-based model to address this shortcoming. We build on state-of-the-art agent-based simulation models but replace the way contacts between susceptible and infected people take place. Firstly, we allow for heterogeneity in the types of contacts (e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we strictly separate the number of contacts from the probabilities that a contact leads to an infection. The number of contacts changes with social distancing policies, the infection probabilities remain invariant. This allows us to model many types of fine-grained policies that cannot easily be incorporated into other models. To validate our model, we show that it can accurately predict the effect of the German November lockdown even if no similar policy has been observed in the time series that were used to estimate the model parameters.

\vspace{1cm}
JEL Classification: C63, I18

Keywords: Covid-19, agent based simulation model, public health measures

\end{abstract}

>>>>

\noindent Governments worldwide are adopting nuanced policy measures to reduce the number of Covid-19 cases while also  minimizing social and economic costs. Epidemiological models have a hard time predicting the effects of such fine-grained policies. We propose a novel simulation-based model to address this shortcoming. We build on state-of-the-art agent-based simulation models but replace the way contacts between susceptible and infected people take place. Firstly, we allow for heterogeneity in the types of contacts (e.g. recurrent or random) and in the infectiousness of each contact type. Secondly, we strictly separate the number of contacts from the probabilities that a contact leads to an infection. The number of contacts changes with social distancing policies, the infection probabilities remain invariant. This allows us to model many types of fine-grained policies that cannot easily be incorporated into other models. To validate our model, we show that it can accurately predict the effect of the German November lockdown even if no similar policy has been observed in the time series that were used to estimate the model parameters.

\vspace{1cm}
\noindent JEL Classification: C63, I18

\noindent Keywords: Covid-19, agent based simulation model, public health measures

### 2020-12-20 paper/calibration_and_estimation.tex
Clean up preamble, fix typos and add section on initial conditions.
Move to Holger's templates and some other - cannot remember - fixes.

Estimates of the incubation period usually give a range from 2 to 12 days. A meta analysis by \cite{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modelled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95 \% CIs) of 1.63 (95 \% CI 1.51 to 1.75) and 0.50 (95 \% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.

@@

\subsection{Duration of Infectiousness}

@@

\subsection{Duration of Symptoms}

@@

The data on how many percent of symptomatic patients will require ICU is pretty thin. We rely on data by the US CDC (\cite{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/master/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project (2020-09-14)}.

From these sources we arrive at the following probabilities of requiring intensive care:

\begin{table}[!ht]
    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
    \label{tab:symptomatic-to-ICU}
    \centering

@@

@@

% Mossong data
We calibrate the parameters for the predicted numbers of contacts from contact diaries of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg \citep{Mossong2008}. Each contact diary contains all contacts an individual had throughout one day, including information on the other person (such as age and gender) and information on the contact. Importantly, for each contact individuals entered of which type the contact (school, leisure, work etc.) was and how frequent the contact with the other person is.

@@

As mentioned in section \ref{sec:matching}, the probability that two individuals are matched can depend on background characteristics. In particular, we allow this probability to depend on age and county of residence. While we do not have good data on geographical assortativeness and just roughly calibrate it such that 60 \% of contacts are within the same county,\footnote{We are working on improving this estimation with mobility data.} we can calibrate the age assortativeness from the same data we use to calibrate the number of contacts.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/assortative_matching_probability_example.png}
    \caption{Distribution of random non-work contacts by age of participants}
    \label{fig:assortativeness}
\end{figure}

@@

\section{Model Validation}

We validate our model in two ways: 1) We look at the in-sample fit over the estimation period. 2) We look at the out-of-sample fit for November. The last one is a challenging test for our model because there was a strong policy change between the estimation period and November. The model convincingly passes both tests.

\subsection{In-Sample Fit}

Despite fitting only four free parameters, the in-sample fit is very good. The best fit is achieved in the largest age groups. This is so mechanically, because we weight the deviations between simulated and observed infection rates by group sizes. The worst fit is achieved for the 80 to 100 years old. The model predicts too few infections for these groups because they have very few contacts in all contact types we have included so far. We plan to address this issue soon by adding another contact type that captures all contacts in the data by \cite{Mossong2008} that we have not included so far. Moreover, we expect an improved model fit when we allow for more free parameters.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/goodness_of_fit_by_age_group}
    \caption{Actual vs. simulated infection and fatality rates}
    \label{fig:goodness_of_fit}
\end{figure}

\subsection{Out-of-Sample Fit}

We can assess the out-of-sample fit by projecting the effect of the lockdown light and comparing it to case numbers until now. It is important to note that this is not just a simple extrapolation of a time trend because the lockdown light only started after the estimation period. The out-of-sample fit can be assessed in Figure~\ref{fig:out-of-sample-fit}.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.7\textwidth]{../figures/out_of_sample_validation}
    \caption{Projected effect of the lockdown-light}
    \label{fig:out-of-sample-fit}
\end{figure}

The model correctly predicts the effect of the lockdown light with reasonable accuracy. In particular, the actual case numbers are between our neutral and pessimistic projection. The plot also shows that ending the lockdown light as planned on November 30 would lead to an explosive growth in case numbers in all scenarios.

>>>>

Estimates of the incubation period usually give a range from 2 to 12 days. A meta analysis by \cite{McAloon2020} comes to the conclusion that ``The incubation period distribution may be modelled with a lognormal distribution with pooled $\mu$ and $\sigma$ parameters (95\% CIs) of 1.63 (95\% CI 1.51 to 1.75) and 0.50 (95\% CI 0.46 to 0.55), respectively.'' For simplicity we discretize this distribution into four bins.

@@

\subsubsection{Duration of Infectiousness}

@@

\subsubsection{Duration of Symptoms}

@@

The data on how many percent of symptomatic patients will require ICU is pretty thin. We rely on data by the US CDC (\cite{Stokes2020}) and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}. Table~\ref{tab:symptomatic-to-ICU} shows our derivations for the probabilities of requiring intensive care per age group.

\begin{table}[tb]
    \caption{Shares of symptomatic patients who will require ICU care by age groups.}
    \label{tab:symptomatic-to-ICU}
    \centering

@@

    \tablenotes{The data is taken from \cite{Stokes2020} and \href{https://github.com/BDI-pathogens/OpenABM-Covid19/blob/572e24ca2dbf7153789a92ad3a27e4c515d0e576/documentation/parameters/parameter_dictionary.md}{the OpenABM-Project}.}

@@

We calibrate the parameters for the predicted numbers of contacts from contact diaries of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg \citep{Mossong2008}. Each contact diary contains all contacts an individual had throughout one day, including information on the other person (such as age and gender) and information on the contact. Importantly, for each contact individuals entered of which type the contact (school, leisure, work etc.) was and how frequent the contact with the other person is.

@@

As mentioned in section \ref{sec:matching}, the probability that two individuals are matched can depend on background characteristics. In particular, we allow this probability to depend on age and county of residence. While we do not have good data on geographical assortativeness and just roughly calibrate it such that 60 \% of contacts are within the same county\footnotemark, we can calibrate the age assortativeness from the same data we use to calibrate the number of contacts.

\footnotetext{We are working on improving this estimation with mobility data.}

\begin{figure}[!tp]
    \centering
    \includegraphics[width=\textwidth]{../figures/assortative_matching_probability_example.png}
    \caption{Distribution of random non-work contacts by age of participants}
    \label{fig:assortativeness}
\end{figure}

@@

### 2020-12-20 paper/introduction.tex
Move to Holger's templates and some other - cannot remember - fixes.

The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions across the world. As a result, many countries were able to reduce the number of positive cases and could relax restrictions imposed on their societies. At the same time, the social and economic costs were enormous. In the second wave, countries are trying to implement more nuanced policies. For example, Germany imposed a ``lockdown-light'' in November. Some businesses are closed and certain leisure activities are prohibited but schools and nurseries stay open.\blfootnote{Gabler and Röhrl are grateful for financial support by the German Research Foundation (DFG) through CRC-TR 224 (Projects C01 and A02, respectively). Gabler is grateful for funding by IZA Institute of Labor Economics.}

@@

\setstretch{1.0}
\begin{enumerate}
    \item At the core of the model, people meet people based on a matching algorithm. We distinguish various types of contacts. Currently, these are households, leisure activities, schools, nurseries and several types of contacts at the workplace. Contact types can be random or recurrent and vary in frequency.
    \item Policies can be implemented as shutting down contact types entirely or partially. The reduction of contacts can be random or systematic, e.g., to allow for essential workers.
    \item Infection probabilities vary across contact types, but are invariant to policies which reduce contacts.
    \item The model achieves a good fit on German infection and fatality rate data even if only the infection probabilities are fit to the data and the remaining parameters are calibrated from the medical literature and datasets on contact frequencies.
    \item High quality Python code for the model is freely available on \href{https://github.com/covid-19-impact-lab/sid}{https://github.com/covid-19-impact-lab/sid}, well documented and very flexible. We are actively looking for researchers who want to use our model for their projects.
\end{enumerate}
\setstretch{1.25}

After a brief literature review we describe our model in detail and validate its in-sample and out-of-sample fit on German data.

@@

More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\cite{Dorn2020a}).

>>>>

The first wave of the Covid-19 pandemic prompted strict lockdowns and restrictions across the world. As a result, many countries were able to reduce the number of positive cases and could relax restrictions imposed on their societies. At the same time, the social and economic costs were enormous. In the second wave, countries are trying to implement more nuanced policies. For example, Germany imposed a ``lockdown-light'' in November. Some businesses are closed and certain leisure activities are prohibited but schools and nurseries stay open.

@@

\begin{enumerate}
    \item At the core of the model, people meet people based on a matching algorithm. We distinguish various types of contacts. Currently, these are households, leisure activities, schools, nurseries and several types of contacts at the workplace. Contact types can be random or recurrent and vary in frequency.
    \item Policies can be implemented as shutting down contact types entirely or partially. The reduction of contacts can be random or systematic, e.g., to allow for essential workers.
    \item Infection probabilities vary across contact types, but are invariant to policies which reduce contacts.
    \item The model achieves a good fit on German infection and fatality rate data even if only the infection probabilities are fit to the data and the remaining parameters are calibrated from the medical literature and datasets on contact frequencies.
    \item High quality Python code for the model is freely available on Github, well documented and very flexible\footnotemark. We are actively looking for researchers who want to use our model for their projects.
\end{enumerate}

\footnotetext{The code can be found under \url{https://github.com/covid-19-impact-lab/sid} and the documentation with tutorials and background information under \url{https://sid-dev.readthedocs.io/}.}

In Section~\ref{sec:literature_review}, we give a short overview of epidemiological models and how they relate to our model. We continue with a detailed model description in Section~\ref{sec:model} and proceed with a description of model parameters and the estimation in Section~\ref{sec:calibration_and_estimation}. The model is validated in Section~\ref{sec:model_validation} by assessing the in-sample fit for reported infections from August to October and by comparing the out-of-sample fit for the period from November until the beginning of Christmas for different lockdown scenarios. We conclude in Section~\ref{sec:conclusion}.

@@

More details about the German model as well as applications to currently discussed German policies can be found in a companion paper (\cite{Dorn2020a}).\comment[id=T]{Needs to be updated with a reference to the new report.}
